Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-08 eurocg/paper.tex
NP Beweis gestrafft

The \emph{intersection graph} $ C = (V,E) $ of $ \Gamma $ is the graph containing a vertex $v_i$ in $ V $ for every $ s_i \in S $ and an edge $ v_i v_j \in E $ between vertices $ v_i, v_j \in V $ if the corresponding edges $ s_i, s_j \in S $ intersect in $ \Gamma $. Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.

Using a standard sweep-line algorithm, computing the intersection graph $C$ of a set of $m$ line segments takes $O(m \log m + I)$ time~\cite{bcko-cgaa-08}, where $I$ is the number of intersections, i.e., the number of edges of $C$.

@@

\section{Hardness of \maxsped for $k\ge3$}\label{sec:hardness}
\soeren{Should the graph in this section too be called G and its drawing $\Gamma$? Also,~\cite{b-sgh-15} claims, paths need to have not only even but equal length, which seems plausible and helps my argument. I'm however not sure if it is actually necessary.}

In this section we close the gap between the known hardness of \maxsped~\cite{b-sgh-15} and the polynomial-time algorithm for 2-plane drawings~\cite{bcgkmn-pped-17}.

@@

\begin{proof}
	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}. 
	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ connected to the three variables appearing in $c_j$. 
	In our gadget-based reduction we create a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED of total ink at least $L$ if and only if $\phi$ is satisfiable.

	All segments

@@

>>>>

The \emph{intersection graph} $ C = (V,E) $ of $ \Gamma $ is the graph containing a vertex $v_i$ in $ V $ for every $ s_i \in S $ and an edge $ v_i v_j \in E $ between vertices $ v_i, v_j \in V $ if the corresponding edges $ s_i, s_j \in S $ intersect in $ \Gamma $. Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.  
%
Using a standard sweep-line algorithm, computing the intersection graph $C$ of a set of $m$ line segments takes $O(m \log m + |E|)$ time~\cite{bcko-cgaa-08}, where $|E|$ is  the number of intersections.

A \emph{symmetric partial edge drawing} (SPED) $D$ of $\Gamma$ draws a fraction $0 < f_s \le 1$ of each edge $s = uv \in S$ by drawing two symmetric edge stubs at $u$ and $v$ of length $f_s \cdot |s|/2$ each. 
No two stubs in $D$ may intersect.
The \emph{ink} or \emph{ink value}  $I(D)$ of a SPED $D$ is the total stub length $I(D) = \sum_{s \in S} f_s |s|$.
In the problem \maxsped the task is to find for a given drawing $\Gamma$ a SPED $D^*$ such that its ink $I(D^*)$ is maximum over all SPEDs.

@@

\section{Hardness of \maxsped for $k\ge3$}\label{sec:hardness}

In this section we close the gap between the known hardness of \maxsped~\cite{b-sgh-15} and the polynomial-time algorithm for 2-plane drawings~\cite{bcgkmn-pped-17} as stated in the following theorem.

@@

\begin{proof}
	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}. %, which reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}. 
	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ either above or below the horizontal line and connected on that side of the line to the three variables appearing in $c_j$. 
	In our gadget-based reduction we create a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
	Figure~\ref{fig:gadget_connection} shows a sketch of the gadgets.

	\begin{figure}
		\centering
		\includegraphics[width=.5\linewidth]{variable_clause_connection}
		\caption{Connection between the clause and variable gadgets, created by chains of an even number of segments.}
		\label{fig:gadget_connection}
	\end{figure}

	All segments in the gadgets are of length 5.
	We use pairs of intersecting segments, alternatingly colored red and green.
	The intersection point of each red-green segment pair is at distance 1 from an endpoint.
	Thus the maximum amount of ink contributed by such a pair is 7 (one full segment of length 5 and the other one with two stubs of length 1 each).

	Each variable gadget is a cycle of segment pairs, with at least one pair for each occurence of the variable in $\phi$, see the three variable gadgets in the bottom of Fig.~\ref{fig:gadget_connection}. 
	Observe that this cycle has exactly two ink-maximal SPEDs: either all red edges are full segments and all green edges are length-1 stubs or vice versa.
	We associate the configuration with full red segments and green stubs with the value \emph{true} and the configuration with full green segments and red stubs with the value \emph{false}.

	For each clause we construct a triple of mutually intersecting segments, see the black triangle at the top of Fig.~\ref{fig:gadget_connection}.
	Again, their intersection points are at distance 1 from the endpoints.
	It is clear that in such a clause triangle at most one of the three segments can be fully drawn, while the stubs of the other two can have length at most 1.
	Hence the maximum amount of ink in a SPED contributed by a clause gadget is 9.

	Finally, we connect variable and clause gadgets in such a way that a clause gadget can contribute its maximum ink value of 9 if and only if the clause is satisfied by the selected truth assignment to the variables.
	For a positive (negative) literal, we create a path of even length between a green (red) edge of the variable gadget and one of the three edges of the clause gadget as shown in Fig.~\ref{fig:gadget_connection}.
	The first edge $s$ of this path intersects the corresponding variable edge $s'$ such that $s'$ is split into a piece of length 2 and a piece of length 3, whereas $s$ is split into a piece of length 1 and a piece of length 4.
	The last edge of the path intersects the corresponding clause edge with the same length ratios.
	The path itself consists of a chain of red-green segment pairs, where, as before, each pair can contribute an ink value of at most~7.

	Since the drawing of $H_\phi$ has polynomial size, the number of segments created in the reduction is polynomial. 
	Further, no segment intersects more than three other segments, so the constructed drawing is 3-plane.

	For the correctness of the reduction, let $L$ be the ink value obtained by counting 7 for each red-green segment pair and 9 for each clause gadget. 
	First assume that $\phi$ has a satisfying truth assignment and put each variable gadget in its corresponding state.
	For each clause, select exactly one literal with value \emph{true} in the satisfying truth assignment. 
	We draw the clause segment that connects to the selected literal as a full segment and the other two as length-1 stubs.
	Since the last segment of the selected literal path must be drawn as length-1 stubs, the only way of having a maximum contribution of that path is by alternating stubs and full segments.
	Hence the first segment of the path must be a full segment.
	But because the variable is in the state that sets the literal to \emph{true}, the intersecting variable segment is drawn as two stubs and the path configuration is valid.
	For the two non-selected literals, we can draw the last segments of their paths as full segments, as well as every segment at an even position, while the segments at odd positions are drawn as stubs. 
	This is compatible with any of the two variable configurations and proves that we can indeed achieve ink value $L$.

	Conversely, assume that we have a SPED with ink $L$. 
	By construction, every red-green segment pair and all clause gadgets must contribute their respective maximum ink value.
	In particular, each variable gadget is either in state \emph{true} or \emph{false}.
	We claim that this truth assignment satisfies $\phi$.
	For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as stubs and all segments at odd positions as full segments. 
	In particular the first segment of the literal path is fully drawn and therefore in a SPED with ink $L$ the intersecting segment of the variable cannot be fully drawn.
	In fact, it must be drawn as length-1 stubs for otherwise the variable gadget would not contribute its maximum ink value.
	But then by construction the selected variable value sets our literal to \emph{true}.
\end{proof}

@@

%%%% OLD TEXT PARTS %%%%%

\begin{proof}
